---
title: "HW5 - Make Data Move"
author: ""
date: "05/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#  Explore global development with R
Today, you will load a filtered gapminder dataset -  with a subset of data on global development from 1952 - 2007 in increments of 5 years - to capture the period between the Second World War and the Global Financial Crisis. 

**Your task: Explore the data and visualise it in both static and animated ways, providing answers and solutions to 7 questions/tasks below.**

## Get the necessary packages
First, start with installing the relevant packages 'tidyverse', 'gganimate', and 'gapminder'.

```{r libraries, echo = FALSE}
# install.packages("gganimate")
# install.packages("gifski")
# install.packages("av")
# install.packages("gapminder")
library(tidyverse)
library(gganimate)
library(gifski)
library(av)
library(gapminder)
```

## Look at the data and tackle the tasks
First, see which specific years are actually represented in the dataset and what variables are being recorded for each country. Note that when you run the cell below, Rmarkdown will give you two results - one for each line - that you can flip between.
```{r}
str(gapminder)
unique(gapminder$year)
head(gapminder)
```

The dataset contains information on each country in the sampled year, its continent, life expectancy, population, and GDP per capita.

Let's plot all the countries in 1952.
```{r 1957}
theme_set(theme_bw())  # set theme to white background for better visibility

ggplot(subset(gapminder, year == 1952), aes(gdpPercap, lifeExp, size = pop)) +
  geom_point() +
  scale_x_log10() 
```

...

We see an interesting spread with an outlier to the right. Answer the following questions, please:


1. *Why does it make sense to have a log10 scale on x axis?*
Using a log10 scale on the x axis allows us a greater overview of the different countries and their gdp. the log10 scale sets the space between 10 and 100 to be the same as between 100 and 1000 and so on. This allows us to make a graph that can showcase the lower numbers of the poorer countries and the higher numbers of the rich countries, without making the graph too big for an easy showcase of data.
2. *Who is the outlier (the richest country in 1952 - far right on x axis)?* 
By using the code below I can conclude that Kuwait is the richest country in 1952

```{r}
gapminder %>% 
  filter(!is.na(gdpPercap)) %>% 
  filter(year == 1952) %>% 
  arrange(desc(gdpPercap))
```


Next, you can generate a similar plot for 2007 and compare the differences
```{r 2007}
ggplot(subset(gapminder, year == 2007), aes(gdpPercap, lifeExp, size = pop)) +
  geom_point(aes(color = continent), alpha = 0.5, width = 0.6, height = 0.4) +
  scale_x_log10() 
```

...

The black bubbles are a bit hard to read, the comparison would be easier with a bit more visual differentiation.

Tasks:

3. *Differentiate the **continents** by color, and fix the axis labels and units to be more legible (**Hint**: the 2.50e+08 is so called "scientific notation", which you might want to eliminate)* 
To add colour to the continents and change the size of the bubbles I enter this command 
"geom_point(aes(color = continent), alpha = 0.5, width = 0.6, height = 0.4)" 

4. *What are the five richest countries in the world in 2007?*
To answer this I use the same code as in question 2, and change the year to 2007. That way I can see that the richest countries are in the following order = Norway, Kuwait, Singapore, USA and Ireland

```{r}
gapminder %>% 
  filter(!is.na(gdpPercap)) %>% 
  filter(year == 2007) %>% 
  arrange(desc(gdpPercap))
```



##  Make it move!

The comparison would be easier if we had the two graphs together, animated. We have a lovely tool in R to do this: the `gganimate` package. Beware that there may be other packages your operating system needs in order to glue interim images into an animation or video. Read the messages when installing the package.

Also, there are *two* ways of animating the gapminder ggplot.

### Option 1: Animate using transition_states() 

The first step is to create the object-to-be-animated
```{r anim1 start}
anim <- ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop)) +
  geom_point(aes(colour = continent)) + labs(x = "GDP per Capita", y = "Life Expectancy") + scale_x_log10()  # convert x to log scale
anim
```

...

This plot collates all the points across time. The next step is to split it into years and animate it. This may take some time, depending on the processing power of your computer (and other things you are asking it to do). Beware that the animation might appear in the bottom right 'Viewer' pane, not in this rmd preview. You need to `knit` the document to get the visual inside an html file.

```{r anim1}
anim + transition_states(year, 
                      transition_length = 1,
                      state_length = 1) + labs(x = "GDP per Capita", y = "Life Expectancy")


```
...

Notice how the animation moves jerkily, 'jumping' from one year to the next 12 times in total. This is a bit clunky, which is why it's good we have another option. 


### Option 2 Animate using transition_time()
This option smoothes the transition between different 'frames', because it interpolates and adds transitional years where there are gaps in the timeseries data.

```{r anim2}
anim2 <- ggplot(gapminder, aes(gdpPercap, lifeExp, size = pop)) +
  geom_point() +
  scale_x_log10() + # convert x to log scale
  transition_time(year)
anim2
```

The much smoother movement in Option 2 will be much more noticeable if you add a title to the chart, that will page through the years corresponding to each frame.

Now, choose one of the animation options and get it to work. You may need to troubleshoot your installation of `gganimate` and other packages 


5. *Can you add a title to one or both of the animations above that will change* 
 *in sync with the animation?* *(**Hint**: search labeling for `transition_states()` and `transition_time()` functions respectively)* 
 
```{r}
ggplot(subset(gapminder, year == 2007), aes(gdpPercap, lifeExp, size = pop)) +
  geom_point(aes(color = continent), alpha = 0.5, width = 0.6, height = 0.4) +
  scale_x_log10() + labs(x = "GDP per Capita", y = "Life Expectancy")
```
 


6. *Can you made the axes' labels and units more readable? Consider expanding the abreviated lables as well as the scientific notation in the legend and x axis to whole numbers.* 
By using the command "labs(x = "GDP per Capita", y = "Life Expectancy")" I expand the abreviated lables of the axes. 

By removing the "scale_x_log10" from the command I expand the X axis to whole numbers. By using "options(scipen = 999)" in the console, I managed to turn the "pop" legend into actual numbers instead of scientific notations. 

```{r}
ggplot(subset(gapminder), aes(gdpPercap, lifeExp, size = pop)) +
  geom_point(aes(color = continent), alpha = 0.5, width = 0.6, height = 0.4) + labs(x = "GDP per Capita", y = "Life Expectancy")
```

 

7. *Come up with a question you want to answer using the gapminder data and write it down. Then, create a data visualisation that answers the question and explain how your visualization answers the question. (Example: you wish to see what was mean life expectancy across the continents in the year you were born versus your parents' birth years). [Hint: if you wish to have more data than is in the filtered gapminder, you can load either the `gapminder_unfiltered` dataset and download more at https://www.gapminder.org/data/ ]*

:)

