---
title: "Project_Suicides"
output: html_document
date: "2023-11-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
```

## Suicide Attempts Graphs

Graphs of suicide attempts in Denmark. Graph over the total attempts and age ranges 15-19, 20-29 and 30-39


Total attempts
```{r Suicide Attempts}
ggplot(data = Suicides2, aes( x = Year, y = Suicide_attempts)) +
  geom_line()
```


15-19
```{r Suicide Attempts 15-19}
ggplot(data = Suicides2, aes( x =Year, y = Suicide_attempts_1519)) + 
  geom_line(colour="red")
```

20-29 
```{r Suicide Attempts 20-29}
ggplot(data = Suicides2, aes( x =Year, y = Suicide_attempts_2029)) + 
  geom_line(colour="orange")
```

30-39
```{r Suicide Attempts 30-39}
ggplot(data = Suicides2, aes( x =Year, y = Suicide_attempts_3039)) + 
  geom_line(colour="blue")
```
Suicide attempts rates compared
```{r Suicide Attempts Rates compared}
ggplot(data = Suicides2, aes( x =Year)) + 
  geom_line(aes( y = Suicide_attempts_2029), colour="orange") +
  geom_line(aes( y = Suicide_attempts_1519), colour="red") + 
  geom_line(aes( y = Suicide_attempts_3039), colour="blue")
```
Red = 15-19
Orange = 20-29
Blue = 30-39

## Suicides


Total suicide rates and for ages 20-29 and 30-39. Because some of the suicides in the 15-19 age group are minors, the data for this group is not available.

Total suicides
```{r Suicides}
ggplot(data = Suicides2, aes( x =Year, y = Suicides)) + 
  geom_line()
```

Suicides 20-29
```{r Suicides 20-29}
ggplot(data = Suicides2, aes( x =Year, y = Suicides_2029)) + 
  geom_line(colour ="orange")
```

Suicides 30-39
```{r Suicides 30-39}
ggplot(data = Suicides2, aes( x =Year, y = Suicides_3039)) + 
  geom_line(colour="blue")
```

Comparison of suicide rates age range 20-29 and 30-39
```{r Suicide Rates comparison}
ggplot(data = Suicides2, aes( x =Year)) + 
  geom_line(aes(y = Suicides_2029), colour="orange") + 
  geom_line(aes(y = Suicides_3039), colour="blue")
```
Orange = 20-29
Blue = 30-39

##Amount of people on antidepressants

```{r People on antidepressants}
ggplot(data = Suicides2, aes( x =Year, y = People_on_antidepressants)) +
  geom_line()
```


Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
